---
title: "Exercise #X"
subtitle: "Fortgeschrittene Statistische Software für NF"
author: "Name (Martrikelnummer), Name (Martrikelnummer), ..."
date: "`r Sys.Date()`"
output: distill::distill_article
---

## General Remarks

-   You can submit your solutions in teams of up to 3 students.
-   Include all your team-member's names and student numbers (Matrikelnummern) in the `authors` field.
-   Please use the exercise template document to work on and submit your results.
-   Use a level 2 heading for each new exercise and answer each subtask next to its bullet points or use a new level 3 heading if you want.
-   Always render the R code for your solutions (`echo=TRUE`) and make sure to include the resulting data in your rendered document.
    -   Make sure to not print more than 10 rows of data (unless specifically instructed to).
-   Always submit both the rendered document(s) as well as your source Rmarkdown or Quarto document. Submit the files separately on moodle, **not** as a zip archive.
-   Submission format is HTML. Other formats will lead to a deduction of points.

## Exercise 1: Initializing git (4 Points)

For this whole exercise sheet we will be tracking all our changes to it in git.

a)  Start by initializing a new R project with git support, called `2025-exeRcise-sheet-3`. If you forgot how to do this, you can follow this [guide](https://malikaihle.github.io/Introduction-RStudio-Git-GitHub/rstudio_project.html).

    done

b)  Commit the files generated by Rstudio.

    done

c)  For all of the following tasks in this exercise sheet we ask you to always commit your changes after finishing each subtask e.g. create a commit after task *1d*, *1e* etc.

> Note: This applies only to answers that have text or code as their answer. If you complete tasks in a different order or forget to commit one, this is no problem. If you change your answers you can just create multiple commits to track the changes.

d)  Name 2 strengths and 2 weaknesses of git. (Don't forget to create a commit after this answer, see *1c*)

    Strengths: Very hard to lose files with git, great for collaboration

    Weaknesses: Struggles with binary (i.e. non-text) files, can be a bit complicated to use (esp. at first)

e)  Knit this exercise sheet. Some new files will automatically be generated when knitting the sheet e.g. the HTML page. Ignore these files, as we only want to track the source files themselves. You can, but don't need to create a `.gitignore` file. Just do not commit these files manually.

## Exercise 2: Putting your Repository on GitHub (3 Points)

For this task you will upload your solution to GitHub.

a)  Create a new repository on GitHub in your account named `exeRcise-sheet-3`. Make sure you create a **public repository** so we are able to see it for grading. Add the link to the repository below:

    <https://github.com/laurenz-n-w/exeRcise-sheet-3>

b)  Push your code to this new repository by copying and executing the snippet on github listed under `…or push an existing repository from the command line`.

    done

c)  Regularly push your latest changes to GitHub again and especially do so when you are finished with this sheet.

## Exercise 3: Pixar Films (4 Points)

Download the `pixar_films` and `public_response` datasets from the GitHub repository and track them in git.

Link: <https://github.com/rfordatascience/tidytuesday/tree/main/data/2025/2025-03-11>

For small datasets like these adding them to git is not a problem.

a)  Load the `pixar_films` dataset into R. Clean the dataset by removing films without a title. Inspect the variable `film_rating`. What are the possible values and what do they mean? Create a factor variable for the film rating. Why is this appropriate?

    ```{r}
    # Load tidyverse packages
    library(tidyverse)

    # Load pixar_films dataset
    data_pixar_films <- read_csv("data/pixar_films.csv")

    # Remove all films without a title
    data_pixar_films <- data_pixar_films %>% 
      filter(!is.na(film))
    ```

    There are three films in the dataset, namely Luca, Turning Red, and Lightyear, without a film rating. Although I have not seen any of these films, by watching the trailers and comparing them to the other films, I believe Luca should be rated G, which means it is suitable for a general audience. The other two films should be rated PG, meaning parental guidance is suggested.

    ```{r}
    # Create factor variable for film rating
    # We use the existing variable, not creating a new one
    data_pixar_films <- data_pixar_films %>% 
      mutate(film_rating = as.factor(film_rating))
    ```

    This is appropriate because there are only two categories: G (General Audience) and PG (Parental Guidance).

b)  Inspect the film titles manually. Which films form a film series? A film series can be identified by a common word in the titles of the films, often in conjunction with a number in the title, e.g. "Despicable Me" and "Despicable Me 2". Create a dataframe which displays a list of the different series with the titles of the films and how many films belong to the series. Output the dataframe.

    ```{r}
    # We inspected the film titles manually and noticed that the only film series are Toy Story, Cars and Incredibles

    # Create series variable
    data_pixar_films <- data_pixar_films %>%
      mutate(series = case_when(
        str_detect(film, "Toy Story") ~ "Toy Story",
        str_detect(film, "Cars") ~ "Cars",
        str_detect(film, "Incredibles") ~ "Incredibles",
        TRUE ~ "None"
      ))

    # Create a dataframe listing the different series with their film titles and the count of films in each series
    different_series <- data_pixar_films %>%
      filter(series != "None") %>% 
      group_by(series) %>%
      summarise(
        films = paste(film, collapse = ", "),
        count = n()
      )

    # Displaying dataframe
    knitr::kable(different_series, caption = "Films per series with titles")
    ```

c)  Load the `public_response` dataframe into R. Convert the `cinema_score` variable into a factor while ensuring the factor levels are defined in ascending order, from the lowest to the highest score. Combine `public_response` with the `pixar_films` dataset using an appropriate merge variable.

    ```{r}
    # Load public_response dataframe
    data_public_response <- read.csv("data/public_response.csv")

    # Convert cinema_score variable into a factor
    data_public_response <- data_public_response %>% 
      mutate(cinema_score = factor(cinema_score, levels = c("A-", "A", "A+"), ordered = TRUE))

    # Check if cinema_score is factor
    is.factor(data_public_response$cinema_score)

    # Check if the factor levels are defined in ascending order, from the lowest to the highest score
    data_public_response %>% 
      select(cinema_score) %>% 
      arrange(cinema_score)

    # Combine public_response and pixar_films dataset
    combined_data <- left_join(data_pixar_films, data_public_response, by = "film")
    ```

d)  Choose one of the variables representing the public response and create a bar plot for the films belonging to a series. Here are the details of the plot:

    -   The film series are represented on the x-axis.
    -   Your chosen public response variable is displayed on the y-axis.
    -   Each film in the series is represented as a separate bar. Bars are grouped by film under their respective series on the x-axis. Order the bars within a series according to the release date of the films.
    -   A title and axis labels for context.

    What do you notice when comparing the scores of the films in a series? Do you see any patterns?

    ```{r}
    # We choose the variable metacritic

    # Creating bar plot for the films belonging to a series
    combined_data %>% 
      filter(series != "None") %>% 
      arrange(series, release_date) %>%  
      mutate(film = factor(film, levels = unique(film))) %>%  
      ggplot(aes(x = series, y = metacritic, fill = film)) +
      geom_col(position = position_dodge()) + 
      labs(
        title = "Public Response (Metacritic) by Film Series",
        x = "Film Series",
        y = "Metacritic Score",
        fill = "Film"
      ) +
      theme_minimal() +
      theme(
      axis.title.x = element_text(size = 10, vjust = -0.5),
      axis.title.y = element_text(size = 10, vjust = 2),
      plot.title = element_text(hjust = 0.5, face = "bold")
    )
    ```

    We observe a clear pattern: the first film in each series consistently earns the highest score, with a noticeable drop between the first and the following installments in the serie.

## Exercise 4: Open Analysis (4 points)

This exercise is a bit more open-ended. You can choose any dataset from [Our World in Data](https://ourworldindata.org/) and analyze it, while determining the research question yourself.

a)  Go to <https://github.com/owid/owid-datasets/tree/master/datasets> and choose a dataset that interests you. You can have a look at <https://ourworldindata.org/> to gather some inspiration.

b)  Download the dataset and track it in git.

    done

c)  Put the name / title of the dataset and a link to it below.

-   Dataset Name: Smoking prevalence and cigarette consumption (IHME, GHDx (2012))
-   Link: [https://raw.githubusercontent.com/owid/owid-datasets/refs/heads/master/datasets/Smoking%20prevalence%20and%20cigarette%20consumption%20(IHME%2C%20GHDx%20(2012))/Smoking%20prevalence%20and%20cigarette%20consumption%20(IHME%2C%20GHDx%20(2012)).csv](https://raw.githubusercontent.com/owid/owid-datasets/refs/heads/master/datasets/Smoking%20prevalence%20and%20cigarette%20consumption%20(IHME%2C%20GHDx%20(2012))/Smoking%20prevalence%20and%20cigarette%20consumption%20(IHME%2C%20GHDx%20(2012)).csv)

d)  Come up with a (research) question you want to answer with the data and briefly explain why you believe this is an interesting question within one sentence. It should be a question that can be answered with the dataset and using R.

    **Which gender showed a greater percentage reduction in the number of daily smokers in Germany, comparing the years 1980 and 2012?**

    This question is interesting because it helps us understand how smoking habits have changed over time for men and women in Germany, especially as knowledge about the health risks of smoking has increased in the last decades.

e)  Use R to answer your chosen question. Please limit your analysis to the functions and techniques we have covered so far in the course. You are **not expected** to use advanced statistical models or external packages which haven't been introduced.

    ```{r}
    # Load the dataset into R
    df_cigarettes <- read_csv("data/cigarette_consumption.csv")

    # Select data for Germany and keep only the relevant columns
    # Rename columns
    df_cigarettes_germany <- df_cigarettes %>% 
      filter(Entity == "Germany") %>% 
      rename(country = Entity, 
             year = Year, 
             female_smokers = `Number of daily female smokers (IHME, GHDx (2012))`, 
             male_smokers = `Number of daily male smokers (IHME, GHDx (2012))`) %>% 
      select(country, year, female_smokers, male_smokers)

    # Preview first 3 rows
    head(df_cigarettes_germany, n = 3) 

    # Filter 1980 and 2012
    df_1980 <- df_cigarettes_germany %>% filter(year == 1980)
    df_2012 <- df_cigarettes_germany %>% filter(year == 2012)

    # Extract values
    male_1980 <- df_1980$male_smokers
    male_2012 <- df_2012$male_smokers
    female_1980 <- df_1980$female_smokers
    female_2012 <- df_2012$female_smokers

    # Calculate percentage reduction
    male_diff <- male_1980 - male_2012
    female_diff <- female_1980 - female_2012

    male_pct <- (male_diff / male_1980) * 100
    female_pct <- (female_diff / female_1980) * 100

    # Print results
    male_pct
    female_pct
    ```

    Between 1980 and 2012, the number of daily smokers in Germany decreased by approximately 21.6% among men and 17.2% among women. This means that men showed a greater percentage reduction in daily smoking during this period.

f)  Create a meaningful plot / figure with the dataset. Make sure to provide a figure caption (via the chunk options / Rmarkdown) and correctly label the figure.

## Final Note

Make sure to push all your commits and changes to GitHub before submitting the exercise sheet.
